---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r gh.url, echo=FALSE}
ghurl <- function(x) {
  x <- unlist(x = strsplit(x = x, split = ','))
  x <- trimws(x = x)
  x <- grep(pattern = "https://github.com", x = x, value = TRUE)[1]
  return(unname(obj = x))
}
```

```{r cran.cannoncial, echo=FALSE}
cran.cannoncial <- "https://cran.r-project.org/package="
```

```{r parse.description, echo=FALSE}
description <- read.dcf(file = "DESCRIPTION")
# Get package name
pkg <- as.vector(x = description[, "Package"])
# Get GitHub URL
github.url <- ghurl(x = description[, "URL"])
# Get depedencies
deps <- as.vector(x = description[, c("Depends", "Imports")])
deps <- unlist(x = strsplit(x = deps, split = "\n"))
deps <- vapply(
  X = deps,
  FUN = function(x) {
    return(unlist(x = strsplit(x = x, split = "(", fixed = TRUE))[1])
  },
  FUN.VALUE = character(length = 1L),
  USE.NAMES = FALSE
)
deps <- trimws(x = deps)
deps <- gsub(pattern = ",$", replacement = "", x = deps)
deps <- deps[deps != "R"]
```

```{r parse.version, echo=FALSE}
pkg.version <- package_version(x = description[1, 'Version'])
stage <- if (pkg.version >= package_version(x = "1.0.0")) {
  "stable"
} else if (pkg.version >= package_version(x = "0.1.0")) {
  "maturing"
} else {
  "experimental"
}
color <- c("experimental" = "orange", "maturing" = "blue", "stable" = "brightgreen")[stage]
```

# `r pkg` v`r pkg.version`

<!-- badges: start -->
[![CRAN/METACRAN](https://img.shields.io/cran/v/`r pkg`)](`r paste0(cran.cannoncial, pkg)`)
[![Lifecycle](https://img.shields.io/badge/lifecycle-`r stage`-`r color`.svg)](`r github.url`)
<!-- badges: end -->

```{r description, echo=FALSE, results='asis'}
cat(gsub(pattern = '\n', replacement = ' ', x = description[, "Description"]))
```

## Installation

You can install the released version of `r pkg` from [CRAN](`r paste0(cran.cannoncial, pkg)`) with:

```{r cran.install, echo=FALSE, results='asis'}
cat(
  "```r",
  paste0("install.packages('", pkg, "')"),
  "```",
  sep = "\n"
)
```

Development versions can be installed from GitHub with:

```{r devel.install, echo=FALSE, results="asis"}
github.base <- unlist(x = strsplit(x = github.url, split = "/"))
github.base <- paste(rev(x = rev(x = github.base)[1:2]), collapse = "/")
cat(
  "```r",
  "if (!requireNamespace('remotes', quietly = TRUE) {",
  "  install.packages('remotes')",
  "}",
  paste0("remotes::install_github('", github.base, "', ref = 'develop')"),
  "```",
  sep = "\n"
)
```
